Synthetic Observations of Extremely High Velocity (EHV) SiO component in young molecular outflows

### 2013-11-26 PAPER/Paola_1/ms_ver2.tex
Added another section in Appendix on value of Gamma

The variation of SiO abundance used for the present model is rather
obtained from empirical values. In case of L1448, \cite{MartinPintado:1992p14309}
have shown that for extreme velocities (obtained from shift in the
spectra) of -40 to -50 km\,s$^{-1}$, the SiO abundance could be as
large as $6\times10^{-7}$, while, it slightly reduces to
$2\times10^{-7}$ for the velocity range of -60 to -50
km\,s$^{-1}$. For the same source, typical {\em{shock}} (or working
surface) velocities of the {\em{molecular bullets}} are estimated to
lie around 80-175 km\,s$^{-1}$ \cite{Dutrey:1997p11185}.
\\
{\color{red}
WE HAVE TWO OPTIONS:\\ 
1. Since the there is a empirical evidence with bulk gas motion. We
can use just the jet velocity obtained from simulations.\\
OR\\
2. We can try to connect the working surface velocities from
\cite{Dutrey:1997p11185} to abundance obtained by
\cite{MartinPintado:1992p14309} and there by use the velocity of the
working surface as an independent variable on the X axis of the
abundance plot.
}

>>>>

The variation of SiO abundance used for the present model is rather
obtained from empirical values. In case of L1448, \cite{MartinPintado:1992p14309}
have shown that for extreme velocities (obtained from shift in the
spectra) of -40 to -50 km\,s$^{-1}$, the SiO abundance could be as
large as $6\times10^{-7}$, while, it slightly increases to
$2\times10^{-6}$ for the velocity range of -60 to -50
km\,s$^{-1}$. For the same source, typical {\em{shock}} (or working
surface) velocities of the {\em{molecular bullets}} are estimated to
lie around 80-175 km\,s$^{-1}$ \cite{Dutrey:1997p11185}.
\\
{\color{red}
WE HAVE TWO OPTIONS:\\ 
1. Since the there is a empirical evidence with bulk gas motion. We
can use just the jet velocity obtained from simulations.\\
OR\\
2. We can try to connect the working surface velocities from
\cite{Dutrey:1997p11185} to abundance obtained by
\cite{MartinPintado:1992p14309} and there by use the velocity of the
working surface as an independent variable on the X axis of the
abundance plot.
}

\section{Treatment of Adiabatic Index $\gamma$}
In dealing with a gas chemistry network which can have both monatomic and diatomic
species, one is immediately presented with the issue of how to treat the adiabatic
index, $\gamma$, of the gas, which directly influences the equation of
state, eq.~\ref{EOS}, and thus the behaviour of the gas.
In standard hydrodynamical codes the gas is generally
assumed to be a monatomic ideal gas, with $\gamma$ = 5/3. 
A diatomic gas, such as H$_{2}$, on the other hand, has a value of $\gamma$
= 1.4 at low temperatures, decreasing to below 1.3
at higher temperatures as the vibrational modes of the hydrogen molecule become
excited. Thus, the value of $\gamma$ for a gas containing a mixture of atomic and
molecular hydrogen requires a temperature-dependent calculation of the $\gamma$
value for H$_2$ and a weighting of the values for the atomic and molecular components
according to their abundance. \cite{Yoshida:2006p13376} have suggested
a prescription of an effective $\gamma$ which takes into account the
corresponding weights of each chemical species and is given by,
\begin{equation}
\frac{1}{\gamma_{\rm eff} - 1} = \frac{1}{\mu}\left[\sum_{i}\frac{X_{i}}{A_{i}}\frac{1}{(\gamma_{i} - 1)}\right]^{-1},
\label{eq:gammaeff}
\end{equation}
where, i is the index for each specie and $\mu$ is the mean molecular
weight. The mass fraction is added in the above equation following the
correction suggested in \cite{OSullivan:2009p10078}.
Additionally, the temperature dependence for example in case of
H$_{2}$ molecule can be obtained using \cite{landau:1980},
\begin{equation}
\frac{1}{\gamma_{H_{2}} - 1} = \frac{1}{2}\left[ 5 +
  2x^{2}\frac{e^{x}}{(e^{x} - 1)^{2}}\right],
\label{eq:gammaH2}
\end{equation}
where the last term with  x = 6100 $K$/T accounts for the vibrational
degree of freedom, with T being the gas temperature.
%

However, for such an implementation in large scale simulations one must first
consider its cost, effectiveness, and even correctness. Although such
a method would indeed yield a more correct $\gamma$ value for the gas
mixture in conditions where the rotational and vibrational modes have
had enough time to reach the level of excitation corresponding to
the temperature of the gas, in cases where the gas has just been
shocked, the method would produce an incorrect $\gamma$ value. 
This is because the gas that passes through shock gets rapidly heated, 
but this energy goes predominantly into the translational degrees of
freedom of the gas. It is only with time that energy goes into the rotational and
vibrational degrees of freedom of the gas through collisions. 
\cite{Flower:2003p11236} maintains that therefore the adiabatic index of
the post-shock gas has close to the ideal monatomic value of 5/3, 
and that in order to calculate correctly its evolution as the gas 
adjusts one would have to model the rotational and vibrational 
levels of the hydrogen molecules in a time-dependent manner.
%

For large-scale simulations present here it is more practical to use
the the standard value of 5/3 which is correct for the post-shock 
and predominantly atomic parts of the flow but incorrect for the
equilibrated parts of the flow with a significant fraction of molecular matter.
Since in this case, as the shock regions and their effect on the molecular hydrogen content of 
the ambient gas are of primary importance here, we opt for the more
relevant, and also simpler, policy of keeping the normal $\gamma$
value of 5/3 for our simulations.
